---
title: "mlm"
author: "Pablo Maldonado"
date: "7/11/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(gridExtra)  
library(mnormt) 
library(lme4) 
library(knitr) 
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(viridis)
```


```{r}
granuloma_data <- read.csv("IHC_ISH_clean_data_WIP 1.csv")
```

```{r}
# Fit Poisson GLMs with offset
glm_granuloma <- glm(rep_tb ~ CD4_count_500 + CD8_count_500 + B220_count_500 + Vaccination + Timepoint + offset(log(total_tb)), data = granuloma_data, family = quasipoisson())


summary(glm_granuloma)

# Coefficients from the model
coefficients <- c(Intercept = -2.70185,
                  CD4_count_500 = -0.19521,
                  CD8_count_500 = 0.22937,
                  B220_count_500 = 0.01623,
                  VaccinationSAL = 0.43675,
                  TimepointD56 = 0.79078)

# Function to calculate percentage change
percent_change <- function(coef) {
  (exp(coef) - 1) * 100
}

# Apply the function to all coefficients (excluding the intercept)
percent_changes <- sapply(coefficients[-1], percent_change)

# Output the percent changes
percent_changes
```
```{r}
colnames(granuloma_data)

# Correlation matrix
cor(granuloma_data[, c("RS_prop", "CD4_count_500")])


cor(granuloma_data[, c("RS_prop", "CD8_count_500")])


cor(granuloma_data[, c("RS_prop", "B220_count_500")])
```


```{r fig.height=3, fig.width=8}
# Plot for B220_percent
ggplot(granuloma_data, aes(x = mouse, y = B220_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of B220_percent") +
  xlab("Mouse") +
  ylab("B220_percent") +
  scale_color_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for CD4_percent
ggplot(granuloma_data, aes(x = mouse, y = CD4_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of CD4_percent") +
  xlab("Mouse") +
  ylab("CD4_percent") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for CD8_percent
ggplot(granuloma_data, aes(x = mouse, y = CD8_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of CD8_percent") +
  xlab("Mouse") +
  ylab("CD8_percent") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for Other_percent
ggplot(granuloma_data, aes(x = mouse, y = Other_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of Other_percent") +
  xlab("Mouse") +
  ylab("Other_percent") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for RS_prop
ggplot(granuloma_data, aes(x = mouse, y = RS_prop, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of RS_prop") +
  xlab("Mouse") +
  ylab("RS_prop") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for RS_index
ggplot(granuloma_data, aes(x = mouse, y = RS_index, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of RS_index") +
  xlab("Mouse") +
  ylab("RS_index") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_rep_tb
ggplot(granuloma_data, aes(x = mouse, y = log_rep_tb, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_rep_tb") +
  xlab("Mouse") +
  ylab("log_rep_tb") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_nonrep_tb
ggplot(granuloma_data, aes(x = mouse, y = log_nonrep_tb, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_nonrep_tb") +
  xlab("Mouse") +
  ylab("log_nonrep_tb") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_total_tb
ggplot(granuloma_data, aes(x = mouse, y = log_total_tb, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_total_tb") +
  xlab("Mouse") +
  ylab("log_total_tb") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_RS_index
ggplot(granuloma_data, aes(x = mouse, y = log_RS_index, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_RS_index") +
  xlab("Mouse") +
  ylab("log_RS_index") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_RS_prop
ggplot(granuloma_data, aes(x = mouse, y = log_RS_prop, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_RS_prop") +
  xlab("Mouse") +
  ylab("log_RS_prop") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = mouse, y = (rep_tb/total_tb), color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of Prop of Replicating Mtb") +
  xlab("Mouse") +
  ylab("Prop of Replicating Mtb") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Cell Percentges vs Proportion of Replicating TB
```{r}
#GOOD
ggplot(granuloma_data, aes(x = Other_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("Other_percent vs Prop of Replicating Mtb") +
  xlab("Other_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggplot(granuloma_data, aes(x = B220_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("B220_percent vs Prop of Replicating Mtb") +
  xlab("B220_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = CD8_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD8_percent vs Prop of Replicating Mtb") +
  xlab("CD8_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = CD4_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Each point is a lesion, may come from the same mouse

```{r}
ggplot(granuloma_data, aes(x = CD8_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ Timepoint) +
  ggtitle("CD8_percent vs Prop of Replicating Mtb") +
  xlab("CD8_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = B220_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ Timepoint) +
  ggtitle("B220_percent vs Prop of Replicating Mtb") +
  xlab("B220_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = Other_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ Timepoint) +
  ggtitle("Other_percent vs Prop of Replicating Mtb") +
  xlab("Other_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = CD4_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ Timepoint) +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Fecteted by mouse, each dot is a lesion within that mouse
```{r}
#GOOD

ggplot(granuloma_data, aes(x = CD4_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = CD8_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD8_percent vs Prop of Replicating Mtb") +
  xlab("CD8_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = B220_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("B220_percent vs Prop of Replicating Mtb") +
  xlab("B220_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = Other_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("Other_percent vs Prop of Replicating Mtb") +
  xlab("Other_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Modeling CD4 ONLY

# Lesion nested within mouse
```{r}
model_nested <- glmer(rep_tb ~ CD4_prop + (1 | mouse/lesion), data = granuloma_data, family = poisson())
model_nested <- glmer(rep_tb ~ CD4_prop + (1 | mouse) + (1 | mouse:lesion), data = granuloma_data, family = poisson()) # Same model as 1st
model_nested <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + (1 | mouse:lesion), data = granuloma_data, family = poisson())


# # Fit the two models
# model1 <- glmer(rep_tb ~ CD4_prop + (1 | mouse/lesion), data = granuloma_data, family = poisson())
# model2 <- glmer(rep_tb ~ CD4_prop + (1 | mouse), data = granuloma_data, family = poisson())
# 
# # Perform the likelihood ratio test
# anova(model1, model2)

# Compute the dispersion statistic
dispersion_statistic <- sum(residuals(model_nested, type = "pearson")^2) / df.residual(model_nested)
dispersion_statistic


# Calculate Pearson residuals
pearson_residuals <- residuals(model_nested, type = "pearson")

# Compute the dispersion statistic
P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_nested) - length(fixef(model_nested)))

P_dispersion_statistic
```

# Lesion nested within mouse + offset to account for total Mtb
```{r}

granuloma_data <- granuloma_data %>%
  mutate(unique_lesion = rownames(granuloma_data))
# # 
# # model_nested_offset <- glmer(rep_tb ~ CD4_prop + (1 | mouse/unique_lesion) + offset(log(total_tb)), data = granuloma_data, family = poisson())
# 
# # model_nested_offset <- glmer(rep_tb ~ CD4_prop + (1 | mouse) + offset(log(total_tb)), data = granuloma_data, family = poisson())
# # 
# model_nested_offset <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + (1 | mouse:unique_lesion) + offset(log(total_tb)), data = granuloma_data, family = poisson())
# 
# summary(model_nested_offset)
# 
# 
# # Compute the dispersion statistic
# dispersion_statistic <- sum(residuals(model_nested_offset , type = "pearson")^2) / df.residual(model_nested_offset )
# dispersion_statistic
# 
# # Calculate Pearson residuals
# pearson_residuals <- residuals(model_nested_offset, type = "pearson")
# 
# # Compute the dispersion statistic
# P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_nested_offset) - length(fixef(model_nested_offset)))
# 
# P_dispersion_statistic
```

# Model accounting for mouse variance
```{r}
# model_mouse <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse), data = granuloma_data, family = poisson())
# 
# 
# summary(model_mouse)
# 
# 
# # Compute the dispersion statistic
# dispersion_statistic <- sum(residuals(model_mouse, type = "pearson")^2) / df.residual(model_mouse)
# dispersion_statistic
# 
# # Calculate Pearson residuals
# pearson_residuals <- residuals(model_mouse, type = "pearson")
# 
# # Compute the dispersion statistic
# P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_mouse) - length(fixef(model_mouse)))
# 
# P_dispersion_statistic
```


# Model accounting for mouse variance + offset of total Mtb
```{r}
# # Fitting the model with the offset
# model_mouse_offset <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + offset(log(total_tb)), 
#                            data = granuloma_data, 
#                            family = poisson())
# 
# # Summary of the model
# summary(model_mouse_offset)
# 
# 
# dispersion_statistic <- sum(residuals(model_mouse_offset, type = "pearson")^2) / df.residual(model_mouse_offset)
# dispersion_statistic
# 
# # Calculate Pearson residuals
# pearson_residuals <- residuals(model_mouse_offset, type = "pearson")
# 
# # Compute the dispersion statistic
# P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_mouse_offset) - length(fixef(model_mouse_offset)))
# 
# P_dispersion_statistic
```


# Negative binomial model accounting for mouse variance + offset of total Mtb
```{r}
# Fit the negative binomial model
# #model_nb <- glmer.nb(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + offset(log(total_tb)), data = granuloma_data)
# # Summary of the model
# summary(model_nb)
# 
# dispersion_statistic <- sum(residuals(model_nb, type = "pearson")^2) / df.residual(model_nb)
# dispersion_statistic
# 
# # Calculate Pearson residuals
# pearson_residuals <- residuals(model_nb, type = "pearson")
# 
# # Compute the dispersion statistic
# P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_nb) - length(fixef(model_nb)))
# 
# P_dispersion_statistic

#https://stackoverflow.com/questions/77564725/nesting-a-random-factor-in-glmm
#https://stats.stackexchange.com/questions/228800/crossed-vs-nested-random-effects-how-do-they-differ-and-how-are-they-specified
```


# Lesion nested within mouse
```{r}
nb <- glmer.nb(rep_tb ~ CD4_count_500 + CD8_count_500 + B220_count_500 + Vaccination + Timepoint + (1 | mouse/unique_lesion) + offset(log(total_tb)), data = granuloma_data)

summary(nb)


model_mouse_offset <- glmer(rep_tb ~ CD4_count_500 + CD8_count_500 + B220_count_500 + 
                              Vaccination + Timepoint + (1 | mouse) + offset(log(total_tb)) + (1 | unique_lesion), 
                            data = granuloma_data, family = poisson())


summary(model_mouse_offset)

# Coefficients from the model
coefficients <- c(Intercept = -2.75987,
                  CD4_count_500 = -0.09134,
                  CD8_count_500 = 0.06960,
                  B220_count_500 = 0.04841,
                  VaccinationSAL = 0.28365,
                  TimepointD56 = 0.96559)

# Function to calculate percentage change
percent_change <- function(coef) {
  (exp(coef) - 1) * 100
}

# Apply the function to all coefficients (excluding the intercept)
percent_changes <- sapply(coefficients[-1], percent_change)

# Output the percent changes
percent_changes
```

```{r}

# Generate predicted values for interaction plot
granuloma_data$predicted <- predict(model_mouse_offset, type = "response")

# Create interaction plot
ggplot(granuloma_data, aes(x = CD4_count_500, y = predicted, color = Vaccination, linetype = Timepoint)) +
  geom_line() +
  labs(title = "Interaction Plot of CD4 Count, Vaccination, and Timepoint",
       x = "CD4 Count (scaled by 500)",
       y = "Predicted Mtb Replication") +
  theme_minimal()

# Calculate residuals
granuloma_data$residuals <- residuals(model_mouse_offset, type = "pearson")

# Plot residuals
ggplot(granuloma_data, aes(x = predicted, y = residuals)) +
  geom_point() +
  labs(title = "Residuals vs Predicted",
       x = "Predicted Mtb Replication",
       y = "Residuals") +
  theme_minimal()

install.packages("emmeans")
library(emmeans)

# Compute estimated marginal means for interaction
emm <- emmeans(model_mouse_offset, ~ CD4_count_500 * Vaccination * Timepoint)

# Plot interaction
plot(emm, comparisons = TRUE)


```

