---
title: "mlm"
author: "Pablo Maldonado"
date: "7/11/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(MASS)
library(gridExtra)  
library(mnormt) 
library(lme4) 
library(knitr) 
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(viridis)
```


```{r}
granuloma_data <- read.csv("IHC_ISH_clean_data_WIP 1.csv")
```

```{r}

colnames(granuloma_data)

granuloma_data$Timepoint <- factor(granuloma_data$Timepoint, levels = c("D56", "D114"))
granuloma_data$Vaccination <- factor(granuloma_data$Vaccination, levels = c("SAL", "BCG"))

ggplot(granuloma_data, aes(x = B220_percent)) +
  geom_histogram(binwidth = 1, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of B220_percent") +
  xlab("B220_percent") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = CD4_percent)) +
  geom_histogram(binwidth = 1, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of CD4_percent") +
  xlab("CD4_percent") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = CD8_percent)) +
  geom_histogram(binwidth = 1, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of CD8_percent") +
  xlab("CD8_percent") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = Other_percent)) +
  geom_histogram(binwidth = 1, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of Other_percent") +
  xlab("Other_percent") +
  ylab("Frequency") + 
  theme_minimal()

```

```{r}

ggplot(granuloma_data, aes(x = RS_prop)) +
  geom_histogram(binwidth = 0.05, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of RS_prop") +
  xlab("RS_prop") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = RS_index)) +
  geom_histogram(binwidth = 0.5, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of RS_index") +
  xlab("RS_index") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = log_rep_tb)) +
  geom_histogram(binwidth = 0.5, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of log_rep_tb") +
  xlab("log_rep_tb") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = log_nonrep_tb)) +
  geom_histogram(binwidth = 0.5, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of log_nonrep_tb") +
  xlab("log_nonrep_tb") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = log_total_tb)) +
  geom_histogram(binwidth = 0.5, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of log_total_tb") +
  xlab("log_total_tb") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = log_RS_index)) +
  geom_histogram(binwidth = 0.5, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of log_RS_index") +
  xlab("log_RS_index") +
  ylab("Frequency") + 
  theme_minimal()

ggplot(granuloma_data, aes(x = log_RS_prop)) +
  geom_histogram(binwidth = 0.5, fill = "black", color = "white") +
  facet_grid(Vaccination ~ Timepoint) +
  ggtitle("Histogram of log_RS_prop") +
  xlab("log_RS_prop") +
  ylab("Frequency") + 
  theme_minimal()

```

```{r fig.height=3, fig.width=8}
# Plot for B220_percent
ggplot(granuloma_data, aes(x = mouse, y = B220_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of B220_percent") +
  xlab("Mouse") +
  ylab("B220_percent") +
  scale_color_viridis_d() +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for CD4_percent
ggplot(granuloma_data, aes(x = mouse, y = CD4_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of CD4_percent") +
  xlab("Mouse") +
  ylab("CD4_percent") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for CD8_percent
ggplot(granuloma_data, aes(x = mouse, y = CD8_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of CD8_percent") +
  xlab("Mouse") +
  ylab("CD8_percent") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for Other_percent
ggplot(granuloma_data, aes(x = mouse, y = Other_percent, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of Other_percent") +
  xlab("Mouse") +
  ylab("Other_percent") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for RS_prop
ggplot(granuloma_data, aes(x = mouse, y = RS_prop, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of RS_prop") +
  xlab("Mouse") +
  ylab("RS_prop") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for RS_index
ggplot(granuloma_data, aes(x = mouse, y = RS_index, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of RS_index") +
  xlab("Mouse") +
  ylab("RS_index") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_rep_tb
ggplot(granuloma_data, aes(x = mouse, y = log_rep_tb, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_rep_tb") +
  xlab("Mouse") +
  ylab("log_rep_tb") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_nonrep_tb
ggplot(granuloma_data, aes(x = mouse, y = log_nonrep_tb, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_nonrep_tb") +
  xlab("Mouse") +
  ylab("log_nonrep_tb") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_total_tb
ggplot(granuloma_data, aes(x = mouse, y = log_total_tb, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_total_tb") +
  xlab("Mouse") +
  ylab("log_total_tb") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_RS_index
ggplot(granuloma_data, aes(x = mouse, y = log_RS_index, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_RS_index") +
  xlab("Mouse") +
  ylab("log_RS_index") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_RS_prop
ggplot(granuloma_data, aes(x = mouse, y = log_RS_prop, color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of log_RS_prop") +
  xlab("Mouse") +
  ylab("log_RS_prop") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = mouse, y = (rep_tb/total_tb), color = lesion)) +
  geom_jitter(width = 0.2) +
  ggtitle("Dotplot of Prop of Replicating Mtb") +
  xlab("Mouse") +
  ylab("Prop of Replicating Mtb") +
  scale_color_viridis_d() +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r fig.height=5, fig.width=10}
# Plot for B220_percent
ggplot(granuloma_data, aes(x = lesion, y = B220_percent)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of B220_percent") +
  xlab("Lesion") +
  ylab("B220_percent")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for CD4_percent
ggplot(granuloma_data, aes(x = lesion, y = CD4_percent)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of CD4_percent") +
  xlab("Lesion") +
  ylab("CD4_percent") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for CD8_percent
ggplot(granuloma_data, aes(x = lesion, y = CD8_percent)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of CD8_percent") +
  xlab("Lesion") +
  ylab("CD8_percent") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for Other_percent
ggplot(granuloma_data, aes(x = lesion, y = Other_percent)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of Other_percent") +
  xlab("Lesion") +
  ylab("Other_percent") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for RS_prop
ggplot(granuloma_data, aes(x = lesion, y = RS_prop)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of RS_prop") +
  xlab("Lesion") +
  ylab("RS_prop") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for RS_index
ggplot(granuloma_data, aes(x = lesion, y = RS_index)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of RS_index") +
  xlab("Lesion") +
  ylab("RS_index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_rep_tb
ggplot(granuloma_data, aes(x = lesion, y = log_rep_tb)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of log_rep_tb") +
  xlab("Lesion") +
  ylab("log_rep_tb") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_nonrep_tb
ggplot(granuloma_data, aes(x = lesion, y = log_nonrep_tb)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of log_nonrep_tb") +
  xlab("Lesion") +
  ylab("log_nonrep_tb") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_total_tb
ggplot(granuloma_data, aes(x = lesion, y = log_total_tb)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of log_total_tb") +
  xlab("Lesion") +
  ylab("log_total_tb") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_RS_index
ggplot(granuloma_data, aes(x = lesion, y = log_RS_index)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of log_RS_index") +
  xlab("Lesion") +
  ylab("log_RS_index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for log_RS_prop
ggplot(granuloma_data, aes(x = lesion, y = log_RS_prop)) +
  geom_jitter(color = "black", width = 0.2) +
  facet_wrap(~ mouse + Timepoint) +
  ggtitle("Dotplot of log_RS_prop") +
  xlab("Lesion") +
  ylab("log_RS_prop") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
# Cell Percentges vs Proportion of Replicating TB
```{r}
ggplot(granuloma_data, aes(x = Other_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("Other_percent vs Prop of Replicating Mtb") +
  xlab("Other_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggplot(granuloma_data, aes(x = B220_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("B220_percent vs Prop of Replicating Mtb") +
  xlab("B220_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = CD8_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD8_percent vs Prop of Replicating Mtb") +
  xlab("CD8_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data, aes(x = CD4_percent, y = (rep_tb/total_tb), color = lesion)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Each point is a lesion, may come from the same mouse
```{r}

granuloma_data_treatment <- granuloma_data %>%
  mutate(treatment = ifelse(grepl("^BCG", mouse), "BCG", "SAL"))



ggplot(granuloma_data_treatment, aes(x = CD8_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  ggtitle("CD8_percent vs Prop of Replicating Mtb") +
  xlab("CD8_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data_treatment, aes(x = B220_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  ggtitle("B220_percent vs Prop of Replicating Mtb") +
  xlab("B220_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data_treatment, aes(x = Other_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  ggtitle("Other_percent vs Prop of Replicating Mtb") +
  xlab("Other_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data_treatment, aes(x = CD4_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Fecteted by mouse, each dot is a lesion within that mouse
```{r}
  
ggplot(granuloma_data_treatment, aes(x = CD8_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD8_percent vs Prop of Replicating Mtb") +
  xlab("CD8_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data_treatment, aes(x = B220_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("B220_percent vs Prop of Replicating Mtb") +
  xlab("B220_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data_treatment, aes(x = Other_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("Other_percent vs Prop of Replicating Mtb") +
  xlab("Other_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(granuloma_data_treatment, aes(x = CD4_percent, y = (rep_tb/total_tb), color = treatment)) +
  geom_point() +
  facet_wrap(~ mouse) +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Modeling CD4

# Lesion nested within mouse
```{r}
model_nested <- glmer(rep_tb ~ CD4_prop + (1 | mouse/lesion), data = granuloma_data, family = poisson())
model_nested <- glmer(rep_tb ~ CD4_prop + (1 | mouse) + (1 | mouse:lesion), data = granuloma_data, family = poisson()) # Same model as 1st
model_nested <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + (1 | mouse:lesion), data = granuloma_data, family = poisson())

summary(model_nested)


# Compute the dispersion statistic
dispersion_statistic <- sum(residuals(model_nested, type = "pearson")^2) / df.residual(model_nested)
dispersion_statistic


# Calculate Pearson residuals
pearson_residuals <- residuals(model_nested, type = "pearson")

# Compute the dispersion statistic
P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_nested) - length(fixef(model_nested)))

P_dispersion_statistic


# # Fit the two models
# model1 <- glmer(rep_tb ~ CD4_prop + (1 | mouse/lesion), data = granuloma_data, family = poisson())
# model2 <- glmer(rep_tb ~ CD4_prop + (1 | mouse), data = granuloma_data, family = poisson())
# 
# # Perform the likelihood ratio test
# anova(model1, model2)


ggplot(granuloma_data_treatment, aes(x = CD4_percent, y = (rep_tb/total_tb), color = Vaccination)) +
  geom_point() +
  facet_wrap(~ Timepoint) +
  ggtitle("CD4_percent vs Prop of Replicating Mtb") +
  xlab("CD4_percent") +
  ylab("Prop of Replicating TB") +
  scale_color_viridis_d() + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

# Lesion nested within mouse + offset to account for total Mtb
```{r}

granuloma_data <- granuloma_data %>% 
  mutate(unique_lesion = rownames(granuloma_data))

model_nested_offset <- glmer(rep_tb ~ CD4_prop + (1 | mouse/unique_lesion) + offset(log(total_tb)), data = granuloma_data, family = poisson())

model_nested_offset <- glmer(rep_tb ~ CD4_prop + (1 | mouse) + offset(log(total_tb)), data = granuloma_data, family = poisson())

model_nested_offset <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + (1 | mouse:lesion) + offset(log(total_tb)), data = granuloma_data, family = poisson())

summary(model_nested_offset)


# Compute the dispersion statistic
dispersion_statistic <- sum(residuals(model_nested_offset , type = "pearson")^2) / df.residual(model_nested_offset )
dispersion_statistic

# Calculate Pearson residuals
pearson_residuals <- residuals(model_nested_offset, type = "pearson")

# Compute the dispersion statistic
P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_nested_offset) - length(fixef(model_nested_offset)))

P_dispersion_statistic
```

# Model accounting for mouse variance
```{r}
model_mouse <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse), data = granuloma_data, family = poisson())


summary(model_mouse)


# Compute the dispersion statistic
dispersion_statistic <- sum(residuals(model_mouse, type = "pearson")^2) / df.residual(model_mouse)
dispersion_statistic

# Calculate Pearson residuals
pearson_residuals <- residuals(model_mouse, type = "pearson")

# Compute the dispersion statistic
P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_mouse) - length(fixef(model_mouse)))

P_dispersion_statistic
```


# Model accounting for mouse variance + offset of total Mtb
```{r}
# Fitting the model with the offset
model_mouse_offset <- glmer(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + offset(log(total_tb)), 
                           data = granuloma_data, 
                           family = poisson())

# Summary of the model
summary(model_mouse_offset)


dispersion_statistic <- sum(residuals(model_mouse_offset, type = "pearson")^2) / df.residual(model_mouse_offset)
dispersion_statistic

# Calculate Pearson residuals
pearson_residuals <- residuals(model_mouse_offset, type = "pearson")

# Compute the dispersion statistic
P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_mouse_offset) - length(fixef(model_mouse_offset)))

P_dispersion_statistic
```


# Negative binomial model accounting for mouse variance + offset of total Mtb
```{r}
# Fit the negative binomial model
model_nb <- glmer.nb(rep_tb ~ CD4_prop + Vaccination * Timepoint + (1 | mouse) + offset(log(total_tb)), data = granuloma_data)

# Summary of the model
summary(model_nb)

dispersion_statistic <- sum(residuals(model_nb, type = "pearson")^2) / df.residual(model_nb)
dispersion_statistic

# Calculate Pearson residuals
pearson_residuals <- residuals(model_nb, type = "pearson")

# Compute the dispersion statistic
P_dispersion_statistic <- sum(pearson_residuals^2) / (nobs(model_nb) - length(fixef(model_nb)))

P_dispersion_statistic

#https://stackoverflow.com/questions/77564725/nesting-a-random-factor-in-glmm
#https://stats.stackexchange.com/questions/228800/crossed-vs-nested-random-effects-how-do-they-differ-and-how-are-they-specified
```

# Modeling "Other"

# Lesion nested within mouse
```{r}

```

# Lesion nested within mouse + offset to account for total Mtb
```{r}

```

# Model accounting for mouse variance
```{r}

```

# Model accounting for mouse variance + offset of total Mtb
```{r}

```

# Negative binomial model accounting for mouse variance + offset of total Mtb
```{r}

```
